\documentclass{article}

\usepackage{comment}
\usepackage[french]{isodate}

\usepackage{graphicx}
\usepackage{siunitx}
\usepackage{paracol}
\usepackage{amsmath}
\usepackage{ amssymb }
\usepackage[utf8]{inputenc}
\usepackage[bookmarks=true]{hyperref}
\usepackage{bookmark}
\usepackage{enumerate}

\usepackage{mathtools,xparse}

\DeclarePairedDelimiter{\abs}{\lvert}{\rvert}
\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}

%Math typeset and settings
\sisetup{output-decimal-marker = {,}}
\newcommand*{\ft}[1]{_\mathrm{#1}} 
\newcommand*{\dd}{\mathop{}\!\mathrm{d}}
\newcommand*{\tran}{^{\mkern-1.5mu\mathsf{T}}}%transpose of matrix


%Math shortcuts
\newcommand{\vout}{v\ft{out}}

\begin{document}
	\begin{titlepage}
		\begin{center}
			\vspace*{1cm}
			\textbf{Math417}\\
			\text{Mathematical Programming}\\
			\vspace{0.5cm}
			Homework IV
			
			\vspace{1.5cm}
			
			\textbf{Frédéric Boileau}\\
			\vspace{2cm}
			Prof. 
			Tim Hoheisel
			\vfill
			\today
			\thispagestyle{empty}
		\end{center}
	\end{titlepage}
	\newpage
	\pdfbookmark{\contentsname}{Contents}
	\tableofcontents
	\thispagestyle{empty}
	\clearpage
	
	\section{Completing the proof of the f. theorem}
	
	Let $\{J_i\}$ denote the set of subsets of $\{1,...n\}$ with $\abs{J}=m$ and such that $a_j$ with $j\in J_i$ are linearly independent $\forall J_i$
	Also let $A^i$ be the matrix
	\begin{equation*}
	\begin{pmatrix}
	a_j\tran\\.\\.\\.	\end{pmatrix} \text{ with } \quad j \in J_i
	\end{equation*}
	Then there exists a unique point, say $\bar x$ such that $A^i\bar x=b$ because by construction it has full rank. Therefore for all $J_i$ there exists only one associated BFS. Moreover there is only a finite number of $J_i$ so there can only be a finite number of BFS.
	
	\section{Derivation of the dual problem}
	
	Here we have two cases
	\begin{enumerate}[i)]
		\item $v\ge 0$
		\begin{equation*}
			\inf \{v\tran x \,\vert \, x\ge 0 \quad v \ge 0 \} = 0
		\end{equation*}
		\item $v_i < 0$ for some i
		\begin{equation*}
			\inf \{v\tran x \, \vert \, x\ge 0 \} = - \infty
		\end{equation*}
		The first case is quite obvious. The support the second we simply say that we can we make the ith components of x go to infinity and keep the others null. So the infinimum is unbounded.
	\end{enumerate}
	\clearpage
	\section{Dual of the dual is the primal}
	We start with the dual problem as denoted in the course notes.
	\begin{align}
		\max \,b\tran y \qquad A\tran y \leq c\\[2ex]
		&\iff -\min-b\tran y \qquad A\tran y +s = c \qquad s\geq 0\\
		&\iff -\min b\tran (y^- - y^+) \qquad A\tran(y^+ - y^-) + s = c
	\end{align}
	Now we have split y to get restrictions on it and introduced a slack variable which must be positive so we have 
	\begin{equation}
		y = y^+ - y^- \qquad y^+, y^-,  s \geq 0
	\end{equation}
	Now we want to express the last problem in standard form. We construct the new vectors z and  $\hat b$ and the new matrix $\hat A$\\[2ex]
	\begin{equation}
		z = \begin{pmatrix} y^+ \\ y^- \\s \end{pmatrix} \qquad
		\hat A = \begin{pmatrix} A\tran &\vert -A\tran &\vert I_n\end{pmatrix}
		\qquad \hat b \tran = \begin{pmatrix} -b\tran & b\tran  & 0_n \end{pmatrix}
	\end{equation}
	\\[2ex]
	We mention the dimensions of these new objects. 
	\begin{equation}
		\hat A \in \mathbb{R}^{n\times 2m+n} \qquad s \in \mathbb{R}^{2m+n}
		\qquad z \in \mathbb{R}^{2m+n} \qquad \hat b \in \mathbb{R}^{2n}
	\end{equation}
	We can now rexpress (1) as follows 
	\begin{equation*}
		- \min \hat b\tran z \qquad \hat A z = c\\
	\end{equation*}
	And we can now take its dual 
	\begin{align*}
		\mathcal{D} : \qquad -&\max c\tran \xi  \qquad \hat A \tran \xi \leq \hat b\\[2ex]
		\iff &\min  c\tran (-\xi) \qquad \hat A (-\xi) \geq -\hat b
	\end{align*}
	
	Now let's take a look at  the constraints
	\begin{equation}
		\begin{pmatrix}
		A\\
		-A\\
		I_n
		\end{pmatrix}
		-\xi =  \begin{pmatrix}
		b\\
		-b\\
		0_n 
		\end{pmatrix}
	\end{equation}
	Which we can break down into 
	\begin{equation}
		A(-\xi) \geq  b \qquad -A(-\xi) \geq -b \qquad -\xi \geq 0
	\end{equation}
	Now let $x = -\xi \geq 0$. We can see from the above reformulation that the two matrix inequalities yield an equality and so we have the standard form of the primal, namely:
	\begin{equation}
		\min c\tran x \qquad Ax = b \qquad x \geq 0 
	\end{equation}
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
	
\end{document}